
```{=latex}
\appendix
% Appendix TOC: chapter-level entries only (Appendix A/B/C/D), no A.1 / A.1.1 etc.
\addtocontents{toc}{\protect\setcounter{tocdepth}{0}}
% Headings already include A.1 / A.1.1; do not prefix .1 / .1.1
\setcounter{secnumdepth}{-1}
% Variable names (`code`) use Times New Roman, same as body text
\setmonofont{Times New Roman}[Ligatures=TeX]
% Wrap identifiers at _ / - (xurl); same Times face as body (\urlstyle{same})
\newcommand{\vn}{\path}
\newcommand{\fmla}[1]{\parbox[c]{\hsize}{\raggedright #1}}
\newcolumntype{Y}[1]{>{\vspace{0pt}\raggedright\arraybackslash}m{#1}}
```

# Appendix A — Dataset details {-}

```{=latex}
\markboth{Appendix A --- Dataset details}{Appendix A --- Dataset details}
```

## A.1 Variable dictionary {-}

All variables were selected from the literature reviewed in Chapter 2.

### A.1.1 Finance / macro (31) {-}

Daily series enter as the Friday last value. Log returns are
$\log(P_{t+1}/P_t)$, as in Chapter 3, and are not multiplied by 100.

The Literature column cites Chapter 2 sources for inclusion.

**Prices & derived (5)**

```{r dict-prices, echo=FALSE, results='asis'}
if (knitr::is_latex_output()) {
  cat(r"(
\begin{longtable}{@{}
  Y{0.28\linewidth}
  Y{0.26\linewidth}
  Y{0.16\linewidth}
  Y{0.26\linewidth}@{}}
\toprule
Variable & Meaning & Source & Literature \\
\midrule
\endfirsthead
\toprule
Variable & Meaning & Source & Literature \\
\midrule
\endhead
\bottomrule
\endlastfoot
\vn{brent_price} & Brent spot (USD/bbl) & EIA & EIA (2014) \\
\vn{wti_price} & WTI Cushing spot (USD/bbl) & EIA & Hao and Wang (2023) \\
\vn{brent_log_return} & Brent weekly log return & derived & Hao and Wang (2023) \\
\vn{wti_log_return} & WTI weekly log return & derived & Hao and Wang (2023) \\
\vn{brent_wti_spread} & Brent − WTI (USD/bbl) & derived & Alquist, Kilian and Vigfusson (2013) \\
\end{longtable}
)")
} else knitr::asis_output("
| Variable | Meaning | Source | Literature |
| --- | --- | --- | --- |
| `brent_price` | Brent spot (USD/bbl) | EIA | EIA (2014) |
| `wti_price` | WTI Cushing spot (USD/bbl) | EIA | Hao and Wang (2023) |
| `brent_log_return` | Brent weekly log return | derived | Hao and Wang (2023) |
| `wti_log_return` | WTI weekly log return | derived | Hao and Wang (2023) |
| `brent_wti_spread` | Brent − WTI (USD/bbl) | derived | Alquist, Kilian and Vigfusson (2013) |
")
```

**EIA WPSR fundamentals (12)**

Source: EIA *Weekly Petroleum Status Report* for all 12 series.

```{r dict-eia, echo=FALSE, results='asis'}
if (knitr::is_latex_output()) {
  cat(r"(
\begin{longtable}{@{}
  Y{0.28\linewidth}
  Y{0.26\linewidth}
  Y{0.16\linewidth}
  Y{0.26\linewidth}@{}}
\toprule
Variable & Meaning & Unit & Literature \\
\midrule
\endfirsthead
\toprule
Variable & Meaning & Unit & Literature \\
\midrule
\endhead
\bottomrule
\endlastfoot
\vn{crude_stocks_excl_spr} & commercial crude stocks excl. SPR & thousand barrels & Alquist, Kilian and Vigfusson (2013); Hao and Wang (2023) \\
\vn{cushing_stocks} & Cushing crude stocks & thousand barrels & Alquist, Kilian and Vigfusson (2013); Hao and Wang (2023) \\
\vn{crude_production} & U.S. crude production & thousand bbl/d & Kilian (2009) \\
\vn{crude_imports} & crude imports & thousand bbl/d & Kilian (2009) \\
\vn{crude_exports} & crude exports & thousand bbl/d & Kilian (2009); Alquist, Kilian and Vigfusson (2013) \\
\vn{refinery_crude_input} & refinery crude input & thousand bbl/d & Kilian (2009) \\
\vn{refinery_utilisation} & refinery utilisation & \% & Kilian (2009) \\
\vn{gasoline_supplied} & gasoline product supplied & thousand bbl/d & Kilian (2009); Costa et al. (2021) \\
\vn{distillate_supplied} & distillate product supplied & thousand bbl/d & Kilian (2009); Costa et al. (2021) \\
\vn{jet_fuel_supplied} & jet-fuel product supplied & thousand bbl/d & Kilian (2009); Costa et al. (2021) \\
\vn{crude_stocks_change} & weekly change in commercial stocks & thousand barrels & Alquist, Kilian and Vigfusson (2013); Hao and Wang (2023) \\
\vn{cushing_stocks_change} & weekly change in Cushing stocks & thousand barrels & Alquist, Kilian and Vigfusson (2013); Hao and Wang (2023) \\
\end{longtable}
)")
} else knitr::asis_output("
| Variable | Meaning | Unit | Literature |
| --- | --- | --- | --- |
| `crude_stocks_excl_spr` | commercial crude stocks excl. SPR | thousand barrels | Alquist, Kilian and Vigfusson (2013); Hao and Wang (2023) |
| `cushing_stocks` | Cushing crude stocks | thousand barrels | Alquist, Kilian and Vigfusson (2013); Hao and Wang (2023) |
| `crude_production` | U.S. crude production | thousand bbl/d | Kilian (2009) |
| `crude_imports` | crude imports | thousand bbl/d | Kilian (2009) |
| `crude_exports` | crude exports | thousand bbl/d | Kilian (2009); Alquist, Kilian and Vigfusson (2013) |
| `refinery_crude_input` | refinery crude input | thousand bbl/d | Kilian (2009) |
| `refinery_utilisation` | refinery utilisation | % | Kilian (2009) |
| `gasoline_supplied` | gasoline product supplied | thousand bbl/d | Kilian (2009); Costa et al. (2021) |
| `distillate_supplied` | distillate product supplied | thousand bbl/d | Kilian (2009); Costa et al. (2021) |
| `jet_fuel_supplied` | jet-fuel product supplied | thousand bbl/d | Kilian (2009); Costa et al. (2021) |
| `crude_stocks_change` | weekly change in commercial stocks | thousand barrels | Alquist, Kilian and Vigfusson (2013); Hao and Wang (2023) |
| `cushing_stocks_change` | weekly change in Cushing stocks | thousand barrels | Alquist, Kilian and Vigfusson (2013); Hao and Wang (2023) |
")
```

**Macro-financial (5)**

```{r dict-macro, echo=FALSE, results='asis'}
if (knitr::is_latex_output()) {
  cat(r"(
\begin{longtable}{@{}
  Y{0.28\linewidth}
  Y{0.26\linewidth}
  Y{0.16\linewidth}
  Y{0.26\linewidth}@{}}
\toprule
Variable & Meaning & Source & Literature \\
\midrule
\endfirsthead
\toprule
Variable & Meaning & Source & Literature \\
\midrule
\endhead
\bottomrule
\endlastfoot
\vn{vix} & CBOE equity-volatility index & FRED VIXCLS & Costa et al. (2021); Yılmaz and Zehir (2026) \\
\vn{dollar_index} & broad nominal USD index & FRED DTWEXBGS & Costa et al. (2021); Yılmaz and Zehir (2026) \\
\vn{treasury_10y} & 10-year Treasury yield & FRED DGS10 & Costa et al. (2021); Yılmaz and Zehir (2026) \\
\vn{fed_funds_rate} & effective federal funds rate & FRED DFF & Costa et al. (2021); Yılmaz and Zehir (2026) \\
\vn{sp500_log_return} & S\&P 500 weekly log return & Yahoo \nolinkurl{^GSPC} & Costa et al. (2021); Yılmaz and Zehir (2026) \\
\end{longtable}
)")
} else knitr::asis_output("
| Variable | Meaning | Source | Literature |
| --- | --- | --- | --- |
| `vix` | CBOE equity-volatility index | FRED VIXCLS | Costa et al. (2021); Yılmaz and Zehir (2026) |
| `dollar_index` | broad nominal USD index | FRED DTWEXBGS | Costa et al. (2021); Yılmaz and Zehir (2026) |
| `treasury_10y` | 10-year Treasury yield | FRED DGS10 | Costa et al. (2021); Yılmaz and Zehir (2026) |
| `fed_funds_rate` | effective federal funds rate | FRED DFF | Costa et al. (2021); Yılmaz and Zehir (2026) |
| `sp500_log_return` | S&P 500 weekly log return | Yahoo ^GSPC | Costa et al. (2021); Yılmaz and Zehir (2026) |
")
```

**Derived market/macro (9)**

```{r dict-derived, echo=FALSE, results='asis'}
if (knitr::is_latex_output()) {
  cat(r"(
\begin{longtable}{@{}
  Y{0.28\linewidth}
  Y{0.26\linewidth}
  Y{0.20\linewidth}
  Y{0.22\linewidth}@{}}
\toprule
Variable & Meaning & Source & Literature \\
\midrule
\endfirsthead
\toprule
Variable & Meaning & Source & Literature \\
\midrule
\endhead
\bottomrule
\endlastfoot
\vn{ovx} & CBOE crude-oil volatility index & Yahoo \nolinkurl{^OVX} & Kilian (2009); Yılmaz and Zehir (2026) \\
\vn{gpr} & geopolitical-risk index (weekly mean of daily GPRD) & Caldara--Iacoviello & Kilian (2009); Yılmaz and Zehir (2026) \\
\vn{gold_return} & gold weekly log return & FRED / Yahoo GC=F & Costa et al. (2021) \\
\vn{global_econ_activity} & Kilian real economic activity (REA) & Dallas Fed IGREA & Kilian (2009) \\
\vn{nonoil_industrial_commodity} & non-fuel industrial materials price index & IMF PINDUINDEXM (FRED) & Kilian (2009); Baumeister and Kilian (2015) \\
\vn{brent_f1_spot_log_basis} & front-month futures − spot log basis & Yahoo BZ=F vs EIA spot & Alquist, Kilian and Vigfusson (2013) \\
\vn{brent_roll_week} & front-month roll-week dummy \{0,1\} & calendar-derived & Alquist, Kilian and Vigfusson (2013) \\
\vn{cadusd_log_return} & CAD/USD weekly log return & Yahoo CADUSD=X & Costa et al. (2021) \\
\vn{dgs10_change} & weekly change in \vn{treasury_10y} & derived & Costa et al. (2021); Yılmaz and Zehir (2026) \\
\end{longtable}
)")
} else knitr::asis_output("
| Variable | Meaning | Source | Literature |
| --- | --- | --- | --- |
| `ovx` | CBOE crude-oil volatility index | Yahoo ^OVX | Kilian (2009); Yılmaz and Zehir (2026) |
| `gpr` | geopolitical-risk index (weekly mean of daily GPRD) | Caldara–Iacoviello | Kilian (2009); Yılmaz and Zehir (2026) |
| `gold_return` | gold weekly log return | FRED / Yahoo GC=F | Costa et al. (2021) |
| `global_econ_activity` | Kilian real economic activity (REA) | Dallas Fed IGREA | Kilian (2009) |
| `nonoil_industrial_commodity` | non-fuel industrial materials price index | IMF PINDUINDEXM (FRED) | Kilian (2009); Baumeister and Kilian (2015) |
| `brent_f1_spot_log_basis` | front-month futures − spot log basis | Yahoo BZ=F vs EIA spot | Alquist, Kilian and Vigfusson (2013) |
| `brent_roll_week` | front-month roll-week dummy {0,1} | calendar-derived | Alquist, Kilian and Vigfusson (2013) |
| `cadusd_log_return` | CAD/USD weekly log return | Yahoo CADUSD=X | Costa et al. (2021) |
| `dgs10_change` | weekly change in `treasury_10y` | derived | Costa et al. (2021); Yılmaz and Zehir (2026) |
")
```

`brent_f1_spot_log_basis` = $\log(\mathrm{BZ{=}F})-\log(\texttt{brent\_price})$; the futures leg is not back-adjusted, so `brent_roll_week` flags the calendar week of the roll. The basis is a public-data proxy for near-term tightness, not a pure calendar spread.

### A.1.2 Remote sensing {-}

**Flat layout (55 = 5 indices × 11 AOI)**

Naming `{index}_anom_{AOI}`; `anom` = within-site deseasonalised z-score
(expanding, past-only). Raw `level` and staleness/mask columns are not modelled.

```{r dict-rs, echo=FALSE, results='asis'}
if (knitr::is_latex_output()) {
  cat(r"(
\begingroup\small
\begin{longtable}{@{}
  Y{0.09\linewidth}
  Y{0.16\linewidth}
  Y{0.28\linewidth}
  Y{0.18\linewidth}
  Y{0.25\linewidth}@{}}
\toprule
Index & Meaning & Formula & Source & Literature \\
\midrule
\endfirsthead
\toprule
Index & Meaning & Formula & Source & Literature \\
\midrule
\endhead
\bottomrule
\endlastfoot
NDVI & vegetation greenness & (B8-B4)\newline/(B8+B4) & Sentinel-2 SR & land-cover control, not oil demand; contrast Bricongne et al. (2026) NO\textsubscript{2} \\
NDWI & surface water/moisture & (B3-B8)\newline/(B3+B8) & Sentinel-2 SR & exploratory (water-adjacent terminals) \\
NDBI & built-up & (B11-B8)\newline/(B11+B8) & Sentinel-2 SR & Wang et al. (2019); Jung (2026) \\
BSI & bare soil / storage yards & ((B11+B4)\newline-(B8+B2))\newline/((B11+B4)\newline+(B8+B2)) & Sentinel-2 SR & Wang et al. (2019); Jung (2026) \\
NTL & night-time light activity & VIIRS DNB \vn{avg_rad} & VIIRS DNB & Polinov, Bookman and Levin (2022); Jung (2026) \\
\end{longtable}
\endgroup
)")
} else knitr::asis_output("
| Index | Meaning | Formula | Source | Literature |
| --- | --- | --- | --- | --- |
| `NDVI` | vegetation greenness | (B8−B4)/(B8+B4) | Sentinel-2 SR | land-cover control, not oil demand; contrast Bricongne et al. (2026) NO~2~ |
| `NDWI` | surface water/moisture | (B3−B8)/(B3+B8) | Sentinel-2 SR | exploratory (water-adjacent terminals) |
| `NDBI` | built-up | (B11−B8)/(B11+B8) | Sentinel-2 SR | Wang et al. (2019); Jung (2026) |
| `BSI` | bare soil / storage yards | ((B11+B4)−(B8+B2))/((B11+B4)+(B8+B2)) | Sentinel-2 SR | Wang et al. (2019); Jung (2026) |
| `NTL` | night-time light activity | VIIRS DNB `avg_rad` | VIIRS DNB | Polinov, Bookman and Levin (2022); Jung (2026) |
")
```

The Flat pathway uses within-site anomalies, not raw levels: night-time
lights mainly capture cross-sectional scale, and raw radiance is a poor
tanker proxy (Polinov, Bookman and Levin, 2022). NDBI and BSI mark yards
and tank farms, not tank-fill; filling rates as in Wang et al. (2019)
need sub-metre imagery. Cloud-quality fields filter composites only;
Hao and Wang’s (2023) cloud-cover mechanism is not replicated.

**Deep layout (Prithvi embeddings)**

Frozen **Prithvi-EO-2.0-300M** embeddings (1024-d per AOI-month; Szwarcman
et al., 2026; cf. SatMAE, Cong et al., 2022). Prithvi was pretrained on
six-band NASA HLS. This study uses Sentinel-2 Surface Reflectance
Harmonized patches, not HLS. Weights are frozen because the weekly Brent
sample is too small to fine-tune a 300-million-parameter encoder. VIIRS
is Flat-only.

| Step | Detail |
| --- | --- |
| Bands | Sentinel-2 `B2, B3, B4, B8A, B11, B12` in HLS order (blue, green, red, narrow NIR, SWIR1, SWIR2) |
| Standardise | published Prithvi `config.json` per-band mean and standard deviation; nodata (0) set to 1e-4 |
| Resample | bilinear resize to 224 × 224 |
| Embedding | frozen encoder; mean-pool of patch tokens → 1024-d; weights never updated |

### A.1.3 Shipping {-}

**Flat layout (164 columns)**

Naming `gfw_{cp}_{stat}` and `pw_{cp}_{stat}` over 6 chokepoints, plus
cross-chokepoint aggregates and PortWatch port export/import volumes, plus
`sar_{region}_{total,dark,share}`. The main Flat shipping specification uses
all 164 columns.

```{r dict-ship-flat, echo=FALSE, results='asis'}
if (knitr::is_latex_output()) {
  cat(r"(
\begingroup\small
\begin{longtable}{@{}
  Y{0.40\linewidth}
  Y{0.32\linewidth}
  Y{0.24\linewidth}@{}}
\toprule
Family & Meaning & Literature \\
\midrule
\endfirsthead
\toprule
Family & Meaning & Literature \\
\midrule
\endhead
\bottomrule
\endlastfoot
\vn{gfw_{cp}_total_hours} /\newline \vn{total_vessels} /\newline \vn{cargo_hours} & GFW vessel-presence hours / distinct vessels / cargo hours & Arslanalp, Marini and Tumbarello (2019) \\
\vn{gfw_{cp}_bunker_hours} /\newline \vn{other_hours} /\newline \vn{other_share} & bunker / other-vessel presence \& share & Arslanalp, Marini and Tumbarello (2019) \\
\vn{gfw_{cp}_total_hours_mom_pct} /\newline \vn{mean_presence_hours_per_vessel} & month-over-month \%; per-vessel congestion proxy & author-derived; congestion concept in Mi et al. (2022) \\
\vn{gfw_all_total_hours_sum} /\newline \vn{gfw_all_activity_zmean} & cross-chokepoint aggregate (sum / leak-free z-mean) & author-derived from GFW \\
\vn{pw_{cp}_n_tanker} /\newline \vn{n_total} /\newline \vn{capacity_tanker} /\newline \vn{capacity} & PortWatch tanker / all-vessel transit count \& capacity & Arslanalp, Marini and Tumbarello (2019); Arslanalp et al. (2026); Adland, Jia and Strandenes (2017); Yan et al. (2020) \\
\vn{pw_{cp}_tanker_share} /\newline \vn{tanker_cap_share} /\newline \vn{avg_tanker_size} & tanker shares; average tanker DWT & author-derived from PortWatch \\
\vn{pw_{cp}_n_tanker_wow_pct} /\newline \vn{capacity_tanker_4w_ma} & week-over-week \%; 4-week MA & author-derived from PortWatch \\
\vn{pw_all_*} (n\_tanker\_sum, n\_total\_sum, tanker\_share) & cross-chokepoint tanker aggregates & author-derived from PortWatch \\
\vn{pw_tanker_exp_imp_net} /\newline \vn{_asym} /\newline \vn{_log_ratio} /\newline \vn{_4w_ma} & export−import net / asymmetry / log-ratio & author-derived; directional flows in Yan et al. (2020) \\
\vn{pw_exp_hubs_export_vol} /\newline \vn{pw_imp_hubs_import_vol} (+\newline \vn{_wow_pct}) & export/import hub tanker tonnage & Arslanalp et al. (2026); Yan et al. (2020) \\
\vn{sar_{region}_{total,dark,share}} & GFW SAR detections: total / unmatched (dark) / dark share; 17 regions × 3 & Paolo et al. (2024) \\
\end{longtable}
\endgroup
)")
} else knitr::asis_output("
| Family | Meaning | Literature |
| --- | --- | --- |
| `gfw_{cp}_total_hours` / `total_vessels` / `cargo_hours` | GFW vessel-presence hours / distinct vessels / cargo hours | Arslanalp, Marini and Tumbarello (2019) |
| `gfw_{cp}_bunker_hours` / `other_hours` / `other_share` | bunker / other-vessel presence & share | Arslanalp, Marini and Tumbarello (2019) |
| `gfw_{cp}_total_hours_mom_pct` / `mean_presence_hours_per_vessel` | month-over-month %; per-vessel congestion proxy | author-derived; congestion concept in Mi et al. (2022) |
| `gfw_all_total_hours_sum` / `gfw_all_activity_zmean` | cross-chokepoint aggregate (sum / leak-free z-mean) | author-derived from GFW |
| `pw_{cp}_n_tanker` / `n_total` / `capacity_tanker` / `capacity` | PortWatch tanker / all-vessel transit count & capacity | Arslanalp, Marini and Tumbarello (2019); Arslanalp et al. (2026); Adland, Jia and Strandenes (2017); Yan et al. (2020) |
| `pw_{cp}_tanker_share` / `tanker_cap_share` / `avg_tanker_size` | tanker shares; average tanker DWT | author-derived from PortWatch |
| `pw_{cp}_n_tanker_wow_pct` / `capacity_tanker_4w_ma` | week-over-week %; 4-week MA | author-derived from PortWatch |
| `pw_all_*` (n_tanker_sum, n_total_sum, tanker_share) | cross-chokepoint tanker aggregates | author-derived from PortWatch |
| `pw_tanker_exp_imp_net` / `_asym` / `_log_ratio` / `_4w_ma` | export−import net / asymmetry / log-ratio | author-derived; directional flows in Yan et al. (2020) |
| `pw_exp_hubs_export_vol` / `pw_imp_hubs_import_vol` (+ `_wow_pct`) | export/import hub tanker tonnage | Arslanalp et al. (2026); Yan et al. (2020) |
| `sar_{region}_{total,dark,share}` | GFW SAR detections: total / unmatched (dark) / dark share; 17 regions × 3 | Paolo et al. (2024) |
")
```

PortWatch `n_tanker` is a liquid-bulk transit count, not unique ships and
not crude-only; `capacity_tanker` is a DWT transit-capacity proxy, not
loaded barrels. GFW presence is general maritime activity, not oil
volume. Derived shares, week-over-week changes and cross-chokepoint sums
are author-constructed from native PortWatch and GFW fields.

**Deep layout — AOI node features (11 per node)**

Node feature spaces differ by type. Graph edges follow maritime-network
studies (Ouyang et al., 2022; Liang et al., 2022; Zhao et al., 2022).

```{r dict-deep-aoi, echo=FALSE, results='asis'}
if (knitr::is_latex_output()) {
  cat(r"(
\begin{longtable}{@{}
  Y{0.28\linewidth}
  Y{0.28\linewidth}
  Y{0.16\linewidth}
  Y{0.24\linewidth}@{}}
\toprule
Variable & Meaning & Source & Literature \\
\midrule
\endfirsthead
\toprule
Variable & Meaning & Source & Literature \\
\midrule
\endhead
\bottomrule
\endlastfoot
\vn{pw_portcalls_tanker} & tanker port calls (weekly sum) & PortWatch & Mi et al. (2022, 2023); Arslanalp et al. (2026) \\
\vn{pw_portcalls_cargo} & cargo port calls & PortWatch & Arslanalp et al. (2026) \\
\vn{pw_import_tanker} & tanker import tonnage & PortWatch & Arslanalp et al. (2026); Yan et al. (2020) \\
\vn{pw_export_tanker} & tanker export tonnage & PortWatch & Arslanalp et al. (2026); Yan et al. (2020) \\
\vn{gfw_n_visits} & port-visit count & GFW AIS & Mi et al. (2022, 2023) \\
\vn{gfw_dwell_hrs_mean} & mean dwell hours & GFW AIS & Mi et al. (2022, 2023) \\
\vn{gfw_dwell_hrs_median} & median dwell hours & GFW AIS & Mi et al. (2022, 2023) \\
\vn{gfw_self_loops} & same-AOI repeat calls & GFW AIS & author-derived from GFW visits \\
\vn{sar_detections_total} & SAR detections & GFW SAR & Paolo et al. (2024) \\
\vn{sar_detections_dark} & unmatched (dark) detections & GFW SAR & Paolo et al. (2024) \\
\vn{sar_dark_share} & dark / total & GFW SAR & Paolo et al. (2024) \\
\end{longtable}
)")
} else knitr::asis_output("
| Variable | Meaning | Source | Literature |
| --- | --- | --- | --- |
| `pw_portcalls_tanker` | tanker port calls (weekly sum) | PortWatch | Mi et al. (2022, 2023); Arslanalp et al. (2026) |
| `pw_portcalls_cargo` | cargo port calls | PortWatch | Arslanalp et al. (2026) |
| `pw_import_tanker` | tanker import tonnage | PortWatch | Arslanalp et al. (2026); Yan et al. (2020) |
| `pw_export_tanker` | tanker export tonnage | PortWatch | Arslanalp et al. (2026); Yan et al. (2020) |
| `gfw_n_visits` | port-visit count | GFW AIS | Mi et al. (2022, 2023) |
| `gfw_dwell_hrs_mean` | mean dwell hours | GFW AIS | Mi et al. (2022, 2023) |
| `gfw_dwell_hrs_median` | median dwell hours | GFW AIS | Mi et al. (2022, 2023) |
| `gfw_self_loops` | same-AOI repeat calls | GFW AIS | author-derived from GFW visits |
| `sar_detections_total` | SAR detections | GFW SAR | Paolo et al. (2024) |
| `sar_detections_dark` | unmatched (dark) detections | GFW SAR | Paolo et al. (2024) |
| `sar_dark_share` | dark / total | GFW SAR | Paolo et al. (2024) |
")
```

Dwell hours keep only `durationHrs` ≤ 720 h (30 days); longer stays are set to missing.

**Deep layout — chokepoint node features (20 per node)**

Same families as the Flat layout, attached to the 6 chokepoint nodes rather
than flattened.

```{r dict-deep-choke, echo=FALSE, results='asis'}
if (knitr::is_latex_output()) {
  cat(r"(
\begin{longtable}{@{}
  Y{0.16\linewidth}
  Y{0.54\linewidth}
  Y{0.26\linewidth}@{}}
\toprule
Block & Features & Literature \\
\midrule
\endfirsthead
\toprule
Block & Features & Literature \\
\midrule
\endhead
\bottomrule
\endlastfoot
GFW (8) & \vn{total_hours}, \vn{total_vessels}, \vn{cargo_hours}, \vn{bunker_hours}, \vn{other_hours}, \vn{other_share}, \vn{total_hours_mom_pct}, \vn{mean_presence_hours_per_vessel} & same as Flat GFW \\
PortWatch (9) & \vn{n_tanker}, \vn{n_total}, \vn{capacity_tanker}, \vn{capacity}, \vn{tanker_share}, \vn{tanker_cap_share}, \vn{avg_tanker_size}, \vn{n_tanker_wow_pct}, \vn{capacity_tanker_4w_ma} & same as Flat PortWatch \\
SAR (3) & \vn{detections_total}, \vn{detections_dark}, \vn{dark_share} & Paolo et al. (2024) \\
\end{longtable}
)")
} else knitr::asis_output("
| Block | Features | Literature |
| --- | --- | --- |
| GFW (8) | `total_hours`, `total_vessels`, `cargo_hours`, `bunker_hours`, `other_hours`, `other_share`, `total_hours_mom_pct`, `mean_presence_hours_per_vessel` | same as Flat GFW |
| PortWatch (9) | `n_tanker`, `n_total`, `capacity_tanker`, `capacity`, `tanker_share`, `tanker_cap_share`, `avg_tanker_size`, `n_tanker_wow_pct`, `capacity_tanker_4w_ma` | same as Flat PortWatch |
| SAR (3) | `detections_total`, `detections_dark`, `dark_share` | Paolo et al. (2024) |
")
```

Mi et al. (2022, 2023) mainly document oil prices affecting tanker port
calls, so shipping is treated as a potentially bidirectional signal, not
a proven leading indicator. SAR dark-vessel detections correct incomplete
public AIS (Paolo et al., 2024); they are not a sanctions variable.

---

## A.2 AOI and chokepoint node lists {-}

### A.2.1 11 oil-infrastructure AOIs {-}

Fixed node order P001–P011 (graph AOI index 0–10). Flat remote-sensing
features use a circular buffer of 5 km radius at every site. Deep Sentinel-2
patches are square and site-specific: 6.4 km for ports, 5.12 km for
refineries, and 1.6–3.2 km for terminals after visual coverage checks.

```{r aoi-sites, echo=FALSE, results='asis'}
if (knitr::is_latex_output()) {
  cat(r"(
\begingroup
\small
\begin{longtable}{@{}
  Y{0.10\linewidth}
  Y{0.18\linewidth}
  Y{0.24\linewidth}
  Y{0.10\linewidth}
  Y{0.16\linewidth}
  Y{0.18\linewidth}@{}}
\toprule
Site ID & Site name & Country / region & Type & Role & Chokepoint \\
\midrule
\endfirsthead
\toprule
Site ID & Site name & Country / region & Type & Role & Chokepoint \\
\midrule
\endhead
\bottomrule
\endlastfoot
P001 & Rotterdam & Netherlands / Europe & port & pricing / import & Suez · Cape \\
P002 & Fujairah & UAE / Middle~East & terminal & transit / storage & Hormuz \\
P003 & Ras Tanura & Saudi Arabia / Middle~East & terminal & export & Hormuz \\
P004 & Jurong Island & Singapore / Asia & refinery & transit / refining & Malacca \\
P005 & Houston & USA / North~America & port & import / refining & Panama \\
P006 & Ningbo-Zhoushan & China / East~Asia & port & import & Malacca \\
P007 & Jamnagar & India / South~Asia & refinery & refining & Hormuz \\
P008 & Al Basrah Terminal & Iraq / Middle~East & terminal & export & Hormuz \\
P009 & Ulsan & South Korea / East~Asia & refinery & refining & Malacca \\
P010 & Kharg Island & Iran / Middle~East & terminal & export & Hormuz \\
P011 & Yanbu & Saudi Arabia / Middle~East & terminal & export & Suez · Mandeb \\
\end{longtable}

\vspace{0.7em}
\begin{longtable}{@{}
  Y{0.12\linewidth}
  Y{0.22\linewidth}
  Y{0.18\linewidth}
  Y{0.18\linewidth}
  Y{0.14\linewidth}@{}}
\toprule
Site ID & Site name & Lat. & Lon. & Patch \\
\midrule
\endfirsthead
\toprule
Site ID & Site name & Lat. & Lon. & Patch \\
\midrule
\endhead
\bottomrule
\endlastfoot
P001 & Rotterdam & 51.950 & 4.145 & 6.4 km \\
P002 & Fujairah & 25.199 & 56.356 & 3.2 km \\
P003 & Ras Tanura & 26.643 & 50.157 & 2.56 km \\
P004 & Jurong Island & 1.274 & 103.708 & 5.12 km \\
P005 & Houston & 29.736 & -95.100 & 6.4 km \\
P006 & Ningbo-Zhoushan & 29.935 & 121.982 & 6.4 km \\
P007 & Jamnagar & 22.345 & 69.860 & 5.12 km \\
P008 & Al Basrah Terminal & 29.681 & 48.810 & 1.6 km \\
P009 & Ulsan & 35.433 & 129.343 & 5.12 km \\
P010 & Kharg Island & 29.231 & 50.324 & 3.2 km \\
P011 & Yanbu & 23.961 & 38.229 & 3.2 km \\
\end{longtable}
\endgroup
)")
} else {
  knitr::kable(
    data.frame(
      `Site ID` = c("P001", "P002", "P003", "P004", "P005", "P006",
                    "P007", "P008", "P009", "P010", "P011"),
      `Site name` = c("Rotterdam", "Fujairah", "Ras Tanura", "Jurong Island", "Houston",
                      "Ningbo-Zhoushan", "Jamnagar", "Al Basrah Terminal", "Ulsan",
                      "Kharg Island", "Yanbu"),
      `Country / region` = c("Netherlands / Europe", "UAE / Middle East", "Saudi Arabia / Middle East",
                             "Singapore / Asia", "USA / North America", "China / East Asia",
                             "India / South Asia", "Iraq / Middle East", "South Korea / East Asia",
                             "Iran / Middle East", "Saudi Arabia / Middle East"),
      Type = c("port", "terminal", "terminal", "refinery", "port", "port", "refinery",
               "terminal", "refinery", "terminal", "terminal"),
      Role = c("pricing / import", "transit / storage", "export", "transit / refining",
               "import / refining", "import", "refining", "export", "refining", "export", "export"),
      Lat. = c("51.950", "25.199", "26.643", "1.274", "29.736", "29.935", "22.345",
               "29.681", "35.433", "29.231", "23.961"),
      Lon. = c("4.145", "56.356", "50.157", "103.708", "−95.100", "121.982", "69.860",
               "48.810", "129.343", "50.324", "38.229"),
      Buffer = "5 km",
      Patch = c("6.4 km", "3.2 km", "2.56 km", "5.12 km", "6.4 km", "6.4 km", "5.12 km",
                "1.6 km", "5.12 km", "3.2 km", "3.2 km"),
      Chokepoint = c("Suez · Cape", "Hormuz", "Hormuz", "Malacca", "Panama", "Malacca",
                     "Hormuz", "Hormuz", "Malacca", "Hormuz", "Suez · Mandeb"),
      check.names = FALSE
    ),
    format = "html",
    align = "l",
    row.names = FALSE
  )
}
```

The circular Flat buffer is 5 km at every site (see the paragraph above). Patch is the side length of the Deep square image chip centred on the same coordinate. Type and Role are the facility type and functional role. The Chokepoint column lists each site's assigned oil-trade corridor(s). Coordinates are WGS84 decimal degrees.

### A.2.2 6 maritime chokepoints {-}

Fixed node order (graph index 11–16), from EIA World Oil Transit Chokepoints.

| Short code | Chokepoint |
| --- | --- |
| `hormuz` | Strait of Hormuz |
| `suez` | Suez Canal |
| `malacca` | Strait of Malacca |
| `mandeb` | Bab el-Mandeb |
| `panama` | Panama Canal |
| `cape` | Cape of Good Hope |

---

## A.3 Publication-lag table {-}

Every predictor enters the weekly (Friday-ending) matrix only after its
conservative as-of availability. Flat and Deep share the same sources but differ
for shipping. This table is the as-of rule used in the main analysis.

### A.3.1 Flat models {-}

```{r flat-lags, echo=FALSE, results='asis'}
if (knitr::is_latex_output()) {
  cat(r"(
\begin{longtable}{@{}
  Y{0.44\linewidth}
  Y{0.32\linewidth}
  Y{0.20\linewidth}@{}}
\toprule
Source & freq $\rightarrow$ weekly & Lag \\
\midrule
\endfirsthead
\toprule
Source & freq $\rightarrow$ weekly & Lag \\
\midrule
\endhead
\bottomrule
\endlastfoot
Daily finance\newline (Brent, WTI, VIX, DXY, DGS10, DFF, S\&P, gold, OVX, CAD) & daily $\rightarrow$ Friday last & \textbf{0} \\
EIA WPSR fundamentals & weekly $\rightarrow$ Friday & \textbf{+1 w} \\
GPR & daily $\rightarrow$ weekly mean & \textbf{+1 w} \\
Monthly macro (REA, non-oil commodity) & month-end, then carried forward & \textbf{+5 w} \\
Sentinel-2 indices + VIIRS (M2) & monthly as-of & \begin{tabular}[c]{@{}l@{}}\textbf{month-end}\\\textbf{+ 15 d}\end{tabular} \\
PortWatch chokepoint/port flows & daily $\rightarrow$ Friday sum & \textbf{+1 w} \\
GFW monthly presence (49 of 164 Flat shipping columns) & month-end, then carried forward & \textbf{+4 w} \\
GFW SAR dark-vessel (51 of 164 Flat shipping columns) & month-end, then carried forward & \textbf{+4 w} \\
\end{longtable}
)")
} else {
  knitr::kable(
    data.frame(
      Source = c(
        "Daily finance (Brent, WTI, VIX, DXY, DGS10, DFF, S&P, gold, OVX, CAD)",
        "EIA WPSR fundamentals",
        "GPR",
        "Monthly macro (REA, non-oil commodity)",
        "Sentinel-2 indices + VIIRS (M2)",
        "PortWatch chokepoint/port flows",
        "GFW monthly presence (49 of 164 Flat shipping columns)",
        "GFW SAR dark-vessel (51 of 164 Flat shipping columns)"
      ),
      `freq → weekly` = c(
        "daily → Friday last",
        "weekly → Friday",
        "daily → weekly mean",
        "month-end, then carried forward",
        "monthly as-of",
        "daily → Friday sum",
        "month-end, then carried forward",
        "month-end, then carried forward"
      ),
      Lag = c("0", "+1 w", "+1 w", "+5 w",
              "month-end + 15 d", "+1 w", "+4 w", "+4 w"),
      check.names = FALSE
    ),
    format = "html",
    align = "l",
    row.names = FALSE
  )
}
```

EIA series are lagged once at construction and are not lagged again at merge.

### A.3.2 Deep models {-}

Finance and remote sensing use the same as-of dates as A.3.1 (Deep remote
sensing = monthly Prithvi embeddings, also month-end + 15 days). Only the
17-node shipping graph differs.

| Graph stream | Role | Lag |
| --- | --- | --- |
| PortWatch node counts | node features | **+1 w** |
| GFW events / voyages (O-D) | edges + node features | **+2 w** |
| GFW SAR dark-vessel | node features | **+4 w** |
| GFW monthly presence (chokepoint node features) | node features | **+4 w** |

---

## A.4 Shipping graph edge definition {-}

Deep models encode shipping as a **weekly 17-node heterogeneous graph**
(11 AOIs + 6 chokepoints, fixed order). Combined adjacency averages about 66
edges per week.

### A.4.1 Dynamic voyage edges (AOI→AOI) {-}

Directed AOI→AOI edges from GFW voyage counts; edge weight = `n_voyages` for
that week's directed lane (`from ≠ to`; self-loops removed to a node feature).
Different every week; 96 lanes, 106 992 voyages total (top lanes e.g.
Ningbo↔Singapore, Fujairah↔Singapore, Singapore↔Rotterdam). Directionality
verified (`P006→P004 ≠ P004→P006`). Lag +2 w.

### A.4.2 Fixed corridor edges (AOI↔chokepoint) {-}

Fixed corridor edges are undirected (13 edges), specified in advance from
each site's main documented oil-trade corridor rather than inferred from weekly
vessel movements or geographic proximity. Present every week.
Every AOI carries at least one fixed corridor edge: P007 (Jamnagar) is a demand-side
refinery rather than a Gulf export terminal, but its crude slate is dominated by
Persian Gulf loadings, so it is attached to Hormuz on the import side.

| Chokepoint | Linked AOIs |
| --- | --- |
| `hormuz` | P002, P003, P007, P008, P010 |
| `suez` | P001, P011 |
| `malacca` | P004, P006, P009 |
| `mandeb` | P011 |
| `cape` | P001 |
| `panama` | P005 |

---

## A.5 Flat remote-sensing coverage {-}

Site-level rates for Section 3.4.3. Weekly-calendar coverage counts a month
for every later Friday that still uses it, so it is not a count of independent
site–month composites.

| Site | Weekly-calendar S2 anomaly coverage | First Friday with S2 anomaly | Monthly S2 composite completeness | Weekly NTL anomaly |
| --- | --- | --- | --- | --- |
| Houston | 100.0% | 2019-01-04 | 100% | 100% |
| Rotterdam | 100.0% | 2019-01-04 | 100% | 100% |
| Jamnagar | 100.0% | 2019-01-04 | 100% | 100% |
| Al Basrah Terminal, Fujairah, Kharg, Ras Tanura, Yanbu | 100.0% | 2019-01-04 | 100% | 100% |
| Ulsan | 93.4% | 2019-06-21 | 100% | 100% |
| Ningbo-Zhoushan | 89.9% | 2019-09-20 | 100% | 100% |
| Jurong Island | 83.8% | 2020-02-21 | 98.4% | 100% |

The shortfall at Ulsan and Ningbo-Zhoushan is the expanding 12-month history
required to define an anomaly, not missing monthly composites. Jurong Island
combines that warm-up with residual cloud gaps.
